Analytical Predictions of Exoplanet Obliquities Generated by Planet-Disk Interactions

### 2020-02-24 writeups/Su_disk_disp.tex
oops forgot to check in updated writeup

%TODO
% rework 3->2->1 ??? part, probably also the 3 -> 3 part
% Fig 7 caption lol
    \documentclass[
        fleqn,
        usenatbib,
        referee,
    ]{mnras}
    \usepackage{
        amsmath,
        amssymb,
        newtxtext,
        newtxmath,
        graphicx,
        ae, aecompl,
        booktabs,
        caption,
        subcaption,
    }
    \usepackage[T1]{fontenc}
    \captionsetup{compatibility=false}

@@

@@

\pubyear{2019}

@@

Spin states satisfying $\p{\rd{\uv{s}}{t}}_{\rm rot} = 0$ are referred to as
\emph{Cassini States} (CSs). They require that $\uv{s}$, $\uv{l}$, and
$\uv{l}_{\rm d}$ be coplanar. There are either two or four CSs, depending
on the values of $\eta$. They are specified by the obliquity $\theta$ and the
precessional phase of $\uv{s}$ around $\uv{l}$, denoted by $\phi$. Following
the standard convention and nomenclature (see Fig.~\ref{fig:cs_locs}), CSs 1, 3, 4
have $\phi = 0$ and $\theta < 0$, corresponding to $\uv{s}$ and $\uv{l}_{\rm
d}$ being on opposite sides of $\uv{l}$, while CS2 has $\phi = \pi$ and $\theta
> 0$, corresponding to $\uv{s}$ and $\uv{l}_{\rm d}$ being on the same side of
$\uv{l}$. When $\eta < \eta_{\rm c}$, where
\begin{equation}
    \eta_{\rm c} \equiv \p{\sin^{2/3}I + \cos^{2/3}I}^{-3/2},\label{eq:etac}
\end{equation}
all four CSs exist, and when $\eta > \eta_{\rm c}$, only CSs 2, 3 exist. The CS
obliquities as a function of $\eta$ are shown in Fig.~\ref{fig:cs_locs}.
\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{../initial/99_misc/2_cs_locs.png}
    \caption{Cassini state obliquities as a function of $\eta$ for $I =
    5^\circ$. Note that $\theta \in [-\pi, \pi]$ is the traditional definition
    of the polar angle. The vertical dashed line indicates $\eta_{\rm c}$ ($=
    0.766$ for $I = 5^\circ$), where CS1 and CS4 merge and annihilate, and the
    horizontal dashed lines indicate $\theta = I$ and $I - 180^\circ$, the
    asymptotic values for CSs 2 and 3 for $\eta \gg \eta_{\rm
    c}$.}\label{fig:cs_locs}
\end{figure}

@@

When $\eta < \eta_{\rm c}$, CS4 exists and is a saddle point. The two
trajectories originating and ending at CS4 are the only two infinite-period
orbits in the phase space. Together, these two critical trajectories are
referred to as the \emph{separatrix} and divide phase space into three zones. In
Fig.~\ref{fig:eq_1contours}, we show the separatrix, the three zones, and their
relations to the CSs. Note that trajectories in zone II librate about CS2
while those in zones I and III circulate.
\begin{figure*}
    \centering
    \includegraphics[width=0.8\textwidth]{../initial/0_eta/1contours_flip.png}
    \caption{Level curves plot of $\mathcal{H}\p{\phi, \cos \theta}$
    [Eq.~\eqref{eq:H}] for $I = 5^\circ$, where warmer colors denote more
    positive values. The black solid line is the separatrix, which only exists
    for $\eta < \eta_{\rm c} = 0.766$. The three zones (I, II, III), divided by
    the separatrix, are labeled. The Cassini states are denoted by filled
    circles and have the same colors as in Fig.~\ref{fig:cs_locs}. The interior
    of the separatrix, shaded in grey, is formally only defined for $\eta <
    \eta_{\rm c}$, but we may identify the points in phase space that flow into
    zone II when evolved forward in time (decreasing $\eta$ adiabatically); this
    is the shaded region in the top left panel, bounded by the black dotted
    line.}\label{fig:eq_1contours}
\end{figure*}

Since $\p{\phi, \cos \theta}$ form a set of canonical variables, the integral
$\oint \cos \theta\;\mathrm{d}\phi$ along a trajectory is an adiabatic invariant
(see Section~\ref{s:ad}). The unsigned areas $\p{\abs{\int \cos \theta
\;\mathrm{d}\phi}}$ of the three zones (as defined in
Fig.~\ref{fig:eq_1contours}) can be computed analytically
\citep{henrard1987,ward2004I}. Define
\begin{subequations}
    \begin{align}
        z_0 &= \eta\cos I, &
        \chi &= \sqrt{-\frac{\tan^3\theta_4}{\tan I} - 1},\\
        \rho &= \chi \frac{\sin^2 \theta_4\cos \theta_4}{
            \chi^2 \cos^2\theta_4 + 1},&
        T &= 2\chi \frac{\cos \theta_4}{
            \chi^2 \cos^2\theta_4 - 1}.
    \end{align}
\end{subequations}
The areas for $\eta < \eta_{\rm c}$ are given by
\begin{subequations}\label{se:area_ward}
    \begin{align}
        A_{\rm I} &= 2\pi\p{1 - z_0} - \frac{A_2}{2}, \label{eq:A1}\\
        A_{\rm II} &= 8\rho + 4\arctan T - 8z_0 \arctan \frac{1}{\chi},
            \label{eq:A2}\\
        A_{\rm III} &= 2\pi\p{1 + z_0} - \frac{A_2}{2}\label{eq:A3}.
    \end{align}
\end{subequations}
These are plotted as a function of $\eta$ in Fig.~\ref{fig:eq_areas}. Note that
the zones are not formally defined for $\eta > \eta_{\rm c}$ since the
separatrix disappears, but a natural extension exists: evolve an initial phase
space point $p$ under adiabatic decrease of $\eta$ until the separatrix appears
at $\eta = \eta_{\rm c}$, then identify $p$ with the zone it is in at $\eta_{\rm
c}$. Since phase space area is conserved under adiabatic evolution, this
extension implies $A_{\rm i}\p{\eta > \eta_{\rm c}} = A_{\rm i}(\eta_{\rm c})$.
The boundary between these extended zones is denoted by the dashed black line in
the top left panel of Fig.~\ref{fig:eq_1contours}, where otherwise no separatrix
exists.
\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{../initial/99_misc/1_areas.png}
    \caption{The solid lines show the fractional areas of each of the zones
    $A_{\rm i}(\eta) / 4\pi$ as given by Eqs.~\eqref{se:area_ward}. The colored
    dashed lines correspond to small $\eta$ approximations used in
    Appendix~\ref{s:ad_approx}. The colored dashed lines for $\eta > \eta_{\rm
    c}$ are the effective values of $A_{\rm II}, A_{\rm III}$ for $\eta >
    \eta_{\rm c}$, denoting the points that would flow into either area under
    adiabatic decrease of $\eta$ from $\eta > \eta_{\rm c}$. The vertical black
    dashed lines correspond to $\eta = \eta_{\rm c}$ [Eq.\eqref{eq:etac}] and
    the values of $\eta$ for which $A_{\rm II}$ is maximized ($\eta_{\rm \max,
    II}$) and for which $A_{\rm III}$ is minimized ($\eta_{\rm \min, III}$,
    Eq.~\eqref{eq:eta_minIII}). The horizontal black dashed line illustrates the
    fraction of phase space occupied by zone II at separatrix appearance.
    }\label{fig:eq_areas}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{../initial/2_toy2/3_ensemble_05_35.png}
    \caption{Top: Shown are the final spin obliquity $\theta_{\rm f}$ as a
    function of the initial spin-disk misalignment angle $\theta_{\rm sd, i}$
    for systems evolving from initial $\eta_{\rm i} \gg 1$ to $\eta_{\rm f} \ll
    1$. The blue dots are results of numerical calculations
    (Section~\ref{ss:ad_ensemble}), and the colored tracks are semi-analytical
    results (Section~\ref{ss:zone_transitions}). The black dotted lines
    represent analytical approximations valid in the small-$\theta_{\rm sd, i}$
    limit (see Appendix~\ref{s:ad_approx}). Bottom: Shown are the probabilities
    of different outcomes. Where a particular $\theta_{\rm sd, i}$ corresponds
    to multiple tracks, the system evolves probabilistically. The track that a
    particular system evolves along in a numerical simulation can be measured by
    examining its final obliquity. The dots represent the inferred probabilities
    from measured final obliquities in our simulations, while the colored tracks
    denote the semi-analytic probability of the system evolving along each
    track. There are five regimes of $\theta_{\rm sd, i}$ values for which
    different tracks are accessible. Semi-analytical calculations of the
    boundaries of these regimes are denoted via the vertical dashed black lines
    (see Section~\ref{ss:zone_transitions}).}\label{fig:ad_ensemble}
\end{figure}
\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{../initial/2_toy2/3_ensemble_10_35.png}
    \caption{Same as the top panel of Fig.~\ref{fig:ad_ensemble} but for $I =
    10^\circ$ and with fewer annotations.}\label{fig:3_ensemble_10_35}
\end{figure}
\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{../initial/2_toy2/3_ensemble_20_35.png}
    \caption{Same as the top panel of Fig.~\ref{fig:3_ensemble_10_35} but for $I
    = 20^\circ$ and with fewer annotations.}\label{fig:3_ensemble_20_35}
\end{figure}

@@

Returning to the evolution of $\uv{s}$, we can classify trajectories by the
sequence of separatrix encounters. Initially, in the $\eta > \eta_{\rm c}$
regime, only zones II and III exist; as $\eta \to 0$, only zones I and III exist
(see Fig.~\ref{fig:eq_1contours}). There are five distinct evolutionary tracks:
\begin{enumerate}
    \item II $\to$ I (see Fig.~\ref{fig:ad_21} for an example). The spin axis
        $\uv{s}$ initially circulates in zone II (see a), and then starts
        librating about CS2 as $\eta$ decreases (b), enclosing some initial
        phase space area $A_{\rm i}$. This libration continues until the
        separatrix expands (due to decreasing $\eta$) to ``touch'' the
        trajectory (see c), at which $A_{\rm II}(\eta_\star) = A_{\rm i}$.
        As $\hat{s}$ to a
        circulating trajectory in zone I immediately bordering the separatrix,
        it will encompass $-A_{\rm I}(\eta_\star)$ phase space area. The final
        obliquity $\theta_{\rm f}$ is then given by Eq.~\eqref{eq:qfaf}, with
        $A_{\rm f} = -A_{\rm I}\p{\eta_\star}$. An analytical approximation to
        $\theta_{\rm f}$ is derived in Appendix~\ref{s:ad_approx} and is
        \begin{equation}
            \p{\cos \theta_{\rm f}}_{\rm II \to I} \approx
                \p{\frac{\pi \theta_{\rm sd, i}^2}{16}}^2 \cot I
                    + \frac{\theta_{\rm sd, i}^2}{4}.\label{eq:qf_21_approx}
        \end{equation}
        The transition probability is
        \begin{equation}
            \Pr\p{\mathrm{II} \to \mathrm{I}} = -\p{
                \frac{\pdil{A_{\rm I}}{\eta}}{\pdil{A_{\rm II}}{\eta}}}
                    _{\eta = \eta_{\star}}.
        \end{equation}

        This track can only occur when the initial condition begins in zone II,
        requiring $A_{\rm i} < A_{\rm II}(\eta_{\rm c})$, where
        \begin{equation}
            A_{\rm II}\p{\eta_c} = 4\pi\s{1 - \p{1 + \tan^{2/3}I}^{-3/2}},
        \end{equation}
        is given by Eq.~\eqref{eq:A2} evaluated at $\eta = \eta_c$
        \citep{ward2004I}.

        Also, note that since $\pdil{A_{\rm I}}{\eta} < 0$ everywhere, while
        $\pdil{A_{\rm II}}{\eta} > 0$ at all possible $\eta_\star$ for an
        initial condition starting in zone II, this track always has nonzero
        probability.

@@

    \item III $\to$ II $\to$ I (see Fig.~\ref{fig:ad_321}) The first separatrix
        encounter occurs at $\eta_1$ when $A_{\rm I}(\eta_1) + A_{\rm
        II}(\eta_1) = A_{\rm i}$, upon which the trajectory moves into zone II
        enclosing intermediate phase space area $A_{\rm m} = A_{\rm
        II}(\eta_1)$. Examining Fig.~\ref{fig:eq_areas}, there exist values $A$
        for which $A_{\rm II}(\eta) = A$ has multiple solutions, and $\eta_1 >
        \eta_{\rm II, \max}$ as labeled. These are the only values for which the
        III $\to$ II $\to$ I transition is permitted. Thus, as $\eta$ continues
        to decrease, a second $\eta_2$ value exists for which $A_{\rm m} =
        A_{\rm II}(\eta_2)$, upon which the trajectory is ejected to zone I and
        $A_f = -A_{\rm I}(\eta_2)$. The final obliquity again then obeys
        Eq.~\eqref{eq:qfaf}.

        Similar to the III $\to$ I track, the III $\to$ II $\to$ I track can
        only occur over the same $A_{\rm II}(\eta_{\rm c}) < A_{\rm i} < \max
        A_{\rm I} + A_{\rm II}$ interval. The probability of encountering this
        track is set by the initial III $\to$ II transition. It bears noting
        that this probability is only nonnegative for a very small fraction of
        $A_{\rm i}$ values, since it requires $\pd{A_{\rm II}}{\eta}$ and
        $\pd{A_{\rm III}}{\eta}$ to have different signs; this occurs only if
        $A_{\rm II}\p{\eta_{\rm c}} < A_{\rm i} < A_{\rm II, \max}$. Outside of
        these bounds, $\Pr_{\rm III \to II} < 0$ which is interpreted again as a
        forbidden transition.

        Then, once a III $\to$ II transition occurs, the second II $\to$ I
        transition occurs for some $\eta_2$ satisfying $A_{\rm II}(\eta_2) =
        A_{\rm II}(\eta_1), \eta_2 < \eta_1$. Graphical inspection shows that
        $\pd{A_{\rm II}}{\eta}$ has the same sign as $\pd{A_{\rm III}}{\eta}$
        over all possible values, so the second II $\to$ I transition is
        guaranteed, completing the III $\to$ II $\to$ I track.

    \item III $\to$ III --- This track depicts the trivial case where no
        separatrix encounter occurs, and $A$ is constant throughout the
        evolution ($A_{\rm f} = A_{\rm i}$). This requires $A_{\rm i} > \max
        \p{A_{\rm I} + A_{\rm II}}$. In the limit of $\eta_{\rm i} \to \infty$
        and $ \eta_{\rm f} \to 0$ we have $\theta_{\rm f} = \theta_{\rm sd, i}$.

        Note that a small deviation from $\theta = \theta_{\rm sd, i}$ occurs
        because we use finite $\eta_{\rm i}$, as hinted at in
        Section~\ref{sss:a_evo}. This is because for $\theta_{\rm sd, i} \gtrsim
        \pi/2$, trajectories are better described as librating about CS3.
        Nevertheless, we can still approximate $A_{\rm i}$ to good accuracy as
        the solid angle enclosed by libration about CS3 with polar angle $2\pi +
        \theta_3 - \theta_2 - \theta_{\rm sd, i}$. This recovers the correct
        transfer function $\theta_{\rm f}\p{\theta_{\rm sd, i}}$ for the III
        $\to$ III regime, as can be seen visibly in
        Fig.~\ref{fig:3_ensemble_20_35}.
\end{enumerate}
In summary, starting from an initial condition with phase space area $A_{\rm i}$
at $\eta = \eta_{\rm i} \gg 1$, the five evolutionary tracks are:
\begin{enumerate}
    \item $A_{\rm i} \in \s{0, A_{\rm II}\p{\eta_{\rm \min, III}}}$. Both the II
        $\to$ III and the II $\to$ I tracks are possible.

@@

    \item $A_{\rm i} > \max \p{A_{\rm I} + A_{\rm II}}$. Only the III $\to$ III
        track.
\end{enumerate}
In all cases, the corresponding ranges for $\theta_{\rm sd,i}$ can be computed via
Eq.~\eqref{eq:ai_qsd}. The boundaries between these ranges are overplotted in
Fig.~\ref{fig:ad_ensemble}, where they can be seen to agree well with the
numerical results.
\begin{figure}
    \centering
    \begin{subfigure}{\columnwidth}
        \centering
        \includegraphics[width=0.5\textwidth]{../initial/2_toy2/3testo21.png}
    \end{subfigure}
    \begin{subfigure}{\columnwidth}
        \centering
        \includegraphics[width=0.5\textwidth]{../initial/2_toy2/3testo21_subplots.png}
    \end{subfigure}
    \caption{Fiducial simulation following the $A_{\rm II} \to A_{\rm I}$
    transition. An initial $\theta_{\rm sd, i} = 0.3\;\mathrm{rad} \approx
    17.2^\circ$ was used, as well as $\epsilon = 3 \times 10^{-4}$. Top plot
    upper panel: plot of $\cos \theta(t)$ (green) in an example simulation.
    Overlaid are the locations of Cassini State 2 (red), and upper and lower
    bounds on the separatrix (black). Note that the trajectory successfully
    tracks CS2 to an obliquity $\theta \approx 90^\circ$. Light vertical blue
    lines denote times portrayed in bottom plot. Top plot lower panel: plot of
    the enclosed separatrix area obtained by integrating the simulated
    trajectory (blue dots) and adiabatic theory (red line). Even though the
    enclosed area jumps at separatrix crossing, the resultant $A$ is still

>>>>

%TODO
% rework 3->2->1 ??? part, probably also the 3 -> 3 part
% Fig 7 caption lol
    \documentclass[
        fleqn,
        usenatbib,
        % referee,
    ]{mnras}
    \usepackage{
        amsmath,
        amssymb,
        newtxtext,
        newtxmath,
        graphicx,
        ae, aecompl,
        booktabs,
        caption,
        subcaption,
    }
    \usepackage[T1]{fontenc}
    \captionsetup{compatibility=false}

@@

% for i in ../initial/99_misc/2_3vec.png ../initial/99_misc/2_cs_locs.png ../initial/0_eta/1contours_flip.png ../initial/99_misc/1_areas.png ../initial/2_toy2/3_ensemble_05_35.png ../initial/2_toy2/3_ensemble_10_35.png ../initial/2_toy2/3_ensemble_20_35.png ../initial/2_toy2/3testo21.png ../initial/2_toy2/3testo21_subplots.png ../initial/2_toy2/3testo23.png ../initial/2_toy2/3testo23_subplots.png ../initial/2_toy2/3testo31.png ../initial/2_toy2/3testo31_subplots.png ../initial/2_toy2/3testo321.png ../initial/2_toy2/3testo321_subplots.png ../initial/2_toy2/3_ensemble_05_25.png ../initial/2_toy2/3_ensemble_05_15.png ../initial/2_toy2/3testo_nonad.png ../initial/2_toy2/3testo_nonad_subplots.png ../initial/2_toy2/3_ensemble_05_05.png ../initial/2_toy2/3scan.png ../initial/2_toy2/3scan_20.png ../initial/99_misc/2_lambdas.png; do cp $i plots_diskdisp; done

@@

\pubyear{2020}

@@

Spin states satisfying $\p{\rd{\uv{s}}{t}}_{\rm rot} = 0$ are referred to as
\emph{Cassini States} (CSs). They require that $\uv{s}$, $\uv{l}$, and
$\uv{l}_{\rm d}$ be coplanar. There are either two or four CSs, depending
on the values of $\eta$. They are specified by the obliquity $\theta$ and the
precessional phase of $\uv{s}$ around $\uv{l}$, denoted by $\phi$. Following
the standard convention and nomenclature (see Fig.~\ref{fig:cs_locs}), CSs 1, 3, 4
have $\phi = 0$ and $\theta < 0$, corresponding to $\uv{s}$ and $\uv{l}_{\rm
d}$ being on opposite sides of $\uv{l}$, while CS2 has $\phi = \pi$ and $\theta
> 0$, corresponding to $\uv{s}$ and $\uv{l}_{\rm d}$ being on the same side of
$\uv{l}$. When $\eta < \eta_{\rm c}$, where
\begin{equation}
    \eta_{\rm c} \equiv \p{\sin^{2/3}I + \cos^{2/3}I}^{-3/2},\label{eq:etac}
\end{equation}
all four CSs exist, and when $\eta > \eta_{\rm c}$, only CSs 2, 3 exist. The CS
obliquities as a function of $\eta$ are shown in Fig.~\ref{fig:cs_locs}.
\begin{figure*}
    \includegraphics[width=0.45\textwidth]{plots_diskdisp/2_3vec.png}\hfil
    %
    \includegraphics[width=0.45\textwidth]{plots_diskdisp/2_cs_locs.png}
    \caption{Left: Definition of angles in the Cassini state configuration and
    the adopted sign convention for $\theta$. Right: Cassini state obliquities
    as a function of $\eta$ for $I = 5^\circ$, where $\theta \in [-\pi, \pi]$
    is the traditional definition of the polar angle. The vertical dashed line
    indicates $\eta_{\rm c}$ ($= 0.766$ for $I = 5^\circ$), where CS1 and CS4
    merge and annihilate, and the horizontal dashed lines indicate $\theta = I$
    and $I - 180^\circ$, the asymptotic values for CSs 2 and 3 for $\eta \gg
    \eta_{\rm c}$.}\label{fig:cs_locs}
\end{figure*}

@@

When $\eta < \eta_{\rm c}$, CS4 exists and is a saddle point. The two
trajectories originating and ending at CS4 are the only two infinite-period
orbits in the phase space. Together, these two critical trajectories are
referred to as the \emph{separatrix} and divide phase space into three zones. In
Fig.~\ref{fig:eq_1contours}, we show the separatrix, the three zones, and their
relations to the CSs. Note that trajectories in zone II librate about CS2
while those in zones I and III circulate.
\begin{figure*}
    \centering
    \includegraphics[width=0.8\textwidth]{plots_diskdisp/1contours_flip.png}
    \caption{Level curves plot of $\mathcal{H}\p{\phi, \cos \theta}$
    [Eq.~\eqref{eq:H}] for $I = 5^\circ$, where warmer colors denote more
    positive values. The black solid line is the separatrix, which only exists
    for $\eta < \eta_{\rm c} = 0.766$. The three zones (I, II, III), divided by
    the separatrix, are labeled. The Cassini states are denoted by filled
    circles and have the same colors as in Fig.~\ref{fig:cs_locs}. The interior
    of the separatrix, shaded in grey, is formally only defined for $\eta <
    \eta_{\rm c}$, but we may identify the points in phase space that flow into
    zone II when evolved forward in time (decreasing $\eta$ adiabatically); this
    is the shaded region in the top left panel, bounded by the black dotted
    line.}\label{fig:eq_1contours}
\end{figure*}

Since $\p{\phi, \cos \theta}$ form a set of canonical variables, the integral
$\oint \cos \theta\;\mathrm{d}\phi$ along a trajectory is an adiabatic invariant
(see Section~\ref{s:ad}). The unsigned areas $\p{\abs{\int \cos \theta
\;\mathrm{d}\phi}}$ of the three zones (as defined in
Fig.~\ref{fig:eq_1contours}) can be computed analytically
\citep{henrard1987,ward2004I}. Define
\begin{subequations}
    \begin{align}
        z_0 &= \eta\cos I, &
        \chi &= \sqrt{-\frac{\tan^3\theta_4}{\tan I} - 1},\\
        \rho &= \chi \frac{\sin^2 \theta_4\cos \theta_4}{
            \chi^2 \cos^2\theta_4 + 1},&
        T &= 2\chi \frac{\cos \theta_4}{
            \chi^2 \cos^2\theta_4 - 1}.
    \end{align}
\end{subequations}
The areas for $\eta < \eta_{\rm c}$ are given by
\begin{subequations}\label{se:area_ward}
    \begin{align}
        A_{\rm I} &= 2\pi\p{1 - z_0} - \frac{A_2}{2}, \label{eq:A1}\\
        A_{\rm II} &= 8\rho + 4\arctan T - 8z_0 \arctan \frac{1}{\chi},
            \label{eq:A2}\\
        A_{\rm III} &= 2\pi\p{1 + z_0} - \frac{A_2}{2}\label{eq:A3}.
    \end{align}
\end{subequations}
These are plotted as a function of $\eta$ in Fig.~\ref{fig:eq_areas}. Note that
the zones are not formally defined for $\eta > \eta_{\rm c}$ since the
separatrix disappears, but a natural extension exists: evolve an initial phase
space point $p$ under adiabatic decrease of $\eta$ until the separatrix appears
at $\eta = \eta_{\rm c}$, then identify $p$ with the zone it is in at $\eta_{\rm
c}$. Since phase space area is conserved under adiabatic evolution, this
extension implies $A_{\rm i}\p{\eta > \eta_{\rm c}} = A_{\rm i}(\eta_{\rm c})$.
The boundary between these extended zones is denoted by the dashed black line in
the top left panel of Fig.~\ref{fig:eq_1contours}, where otherwise no separatrix
exists.
\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{plots_diskdisp/1_areas.png}
    \caption{The solid lines show the fractional areas of each of the zones
    $A_{\rm i}(\eta) / 4\pi$ as given by Eqs.~\eqref{se:area_ward}. The colored
    dashed lines correspond to small $\eta$ approximations used in
    Appendix~\ref{s:ad_approx}. The colored dashed lines for $\eta > \eta_{\rm
    c}$ are the effective values of $A_{\rm II}, A_{\rm III}$ for $\eta >
    \eta_{\rm c}$, denoting the points that would flow into either area under
    adiabatic decrease of $\eta$ from $\eta > \eta_{\rm c}$. The vertical black
    dashed lines correspond to $\eta = \eta_{\rm c}$ [Eq.\eqref{eq:etac}] and
    the values of $\eta$ for which $A_{\rm II}$ is maximized ($\eta_{\rm \max,
    II}$) and for which $A_{\rm III}$ is minimized ($\eta_{\rm \min, III}$,
    Eq.~\eqref{eq:eta_minIII}). The horizontal black dashed line illustrates the
    fraction of phase space occupied by zone II at separatrix appearance.
    }\label{fig:eq_areas}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{plots_diskdisp/3_ensemble_05_35.png}
    \caption{Top: Shown are the final spin obliquity $\theta_{\rm f}$ as a
    function of the initial spin-disk misalignment angle $\theta_{\rm sd, i}$
    for systems evolving from initial $\eta_{\rm i} \gg 1$ to $\eta_{\rm f} \ll
    1$. The blue dots are results of numerical calculations
    (Section~\ref{ss:ad_ensemble}), and the colored tracks are semi-analytical
    results (Section~\ref{ss:zone_transitions}). The black dotted lines
    represent analytical approximations valid in the small-$\theta_{\rm sd, i}$
    limit (see Appendix~\ref{s:ad_approx}). Bottom: Shown are the probabilities
    of different outcomes. Where a particular $\theta_{\rm sd, i}$ corresponds
    to multiple tracks, the system evolves probabilistically. The track that a
    particular system evolves along in a numerical simulation can be measured by
    examining its final obliquity. The dots represent the inferred probabilities
    from measured final obliquities in our simulations, while the colored tracks
    denote the semi-analytic probability of the system evolving along each
    track. There are five regimes of $\theta_{\rm sd, i}$ values for which
    different tracks are accessible. Semi-analytical calculations of the
    boundaries of these regimes are denoted via the vertical dashed black lines
    (see Section~\ref{ss:zone_transitions}).}\label{fig:ad_ensemble}
\end{figure}
\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{plots_diskdisp/3_ensemble_10_35.png}
    \caption{Same as the top panel of Fig.~\ref{fig:ad_ensemble} but for $I =
    10^\circ$ and with fewer annotations.}\label{fig:3_ensemble_10_35}
\end{figure}
\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{plots_diskdisp/3_ensemble_20_35.png}
    \caption{Same as the top panel of Fig.~\ref{fig:3_ensemble_10_35} but for $I
    = 20^\circ$ and with fewer annotations.}\label{fig:3_ensemble_20_35}
\end{figure}

@@

Returning to the evolution of $\uv{s}$, we can classify trajectories by the
sequence of separatrix encounters. Initially, in the $\eta > \eta_{\rm c}$
regime, only zones II and III exist; as $\eta \to 0$, only zones I and III exist
(see Fig.~\ref{fig:eq_1contours}). There are five distinct evolutionary tracks:
\begin{enumerate}
    \item II $\to$ I (see Fig.~\ref{fig:ad_21} for an example). The spin axis
        $\uv{s}$ initially circulates in zone II (see a), and then starts
        librating about CS2 as $\eta$ decreases (b), enclosing some initial
        phase space area $A_{\rm i}$. This libration continues until the
        separatrix expands (due to decreasing $\eta$) to ``touch'' the
        trajectory (see c), at which $A_{\rm II}(\eta_\star) = A_{\rm i}$.
        As $\hat{s}$ to a
        circulating trajectory in zone I immediately bordering the separatrix,
        it will encompass $-A_{\rm I}(\eta_\star)$ phase space area. The final
        obliquity $\theta_{\rm f}$ is then given by Eq.~\eqref{eq:qfaf}, with
        $A_{\rm f} = -A_{\rm I}\p{\eta_\star}$. An analytical approximation to
        $\theta_{\rm f}$ is derived in Appendix~\ref{s:ad_approx} and is
        \begin{equation}
            \p{\cos \theta_{\rm f}}_{\rm II \to I} \approx
                \p{\frac{\pi \theta_{\rm sd, i}^2}{16}}^2 \cot I
                    + \frac{\theta_{\rm sd, i}^2}{4}.\label{eq:qf_21_approx}
        \end{equation}
        The transition probability is
        \begin{equation}
            \Pr\p{\mathrm{II} \to \mathrm{I}} = -\p{
                \frac{\pdil{A_{\rm I}}{\eta}}{\pdil{A_{\rm II}}{\eta}}}
                    _{\eta = \eta_{\star}}.
        \end{equation}
        This track can only occur when the initial condition begins in zone II,
        requiring $A_{\rm i} < A_{\rm II}(\eta_{\rm c})$, where $A_{\rm
        II}\p{\eta_c}$ is given by Eq.~\eqref{eq:A2} evaluated at $\eta =
        \eta_c$. Note that since $\pdil{A_{\rm I}}{\eta} < 0$ everywhere, while
        $\pdil{A_{\rm II}}{\eta} > 0$ at all possible $\eta_\star$ for an
        initial condition starting in zone II, this track always has nonzero
        probability.

@@

    \item III $\to$ II $\to$ I (see Fig.~\ref{fig:ad_321}) That $A_{\rm II}
        \p{\eta_c}$ is not monotonic is key to the existence of this track.
        Consider a trajectory originating in zone III that first encounters the
        separatrix at $\eta_1$, when $A_{\rm I}(\eta_1) + A_{\rm II}(\eta_1) =
        A_{\rm i}$, such that it transitions into zone II enclosing
        intermediate phase space area $A_{\rm m} = A_{\rm II}(\eta_1)$. Such a
        transition has probability
        \begin{equation}
            \Pr\p{\rm III \to II} = -\p{
                \frac{\pdil{A_{\rm II}}{\eta}}{\pdil{A_{\rm III}}{\eta}}}
                    _{\eta = \eta_1},
        \end{equation}
        which is nonnegative (i.e.\ the transition is permitted) if $\eta_1
        \in [\eta_{\rm \max, II}, \eta_c]$. Equivalently, this requires
        $A_{\rm i} \in \s{A_{\rm II}\p{\eta_{\rm c}} , A_{\rm II, \max}}$.

        Then, as $\eta$ continues to decrease, a second $\eta_2$ value exists
        for which $A_{\rm m} = A_{\rm II}(\eta_2)$, upon which the trajectory
        is ejected to zone I and $A_f = -A_{\rm I}(\eta_2)$. Note that $\eta_2
        < \eta_{\rm \max, II}$ necessarily, as zone II must be shrinking in
        order for the trajectory to be ejected. The final obliquity is again
        given by Eq.~\eqref{eq:qfaf}.

        Graphical inspection of Fig.~\ref{fig:eq_areas} shows that
        $\pdil{A_{\rm II}}{\eta}$ and $\pdil{A_{\rm I}}{\eta}$ have opposite
        signs for $\eta_2 < \eta_{\rm \max, II}$, and therefore the second II
        $\to$ I transition is guaranteed. In summary, the III $\to$ II $\to$ I
        track is permitted so long as the first transition is permitted, or
        $A_{\rm i} \in \s{A_{\rm II}\p{\eta_{\rm c}} , A_{\rm II, \max}}$.

    \item III $\to$ III --- This track is the trivial case where no separatrix
        encounter occurs, and $A$ is constant throughout the evolution ($A_{\rm
        f} = A_{\rm i}$). This requires $A_{\rm i} > \max \p{A_{\rm I} + A_{\rm
        II}}$. In the limit of $\eta_{\rm i} \to \infty$ and $ \eta_{\rm f} \to
        0$ we have $\theta_{\rm f} = \theta_{\rm sd, i}$.

        Note that a small deviation from $\theta = \theta_{\rm sd, i}$ occurs
        because we use finite $\eta_{\rm i}$, as hinted at in
        Section~\ref{sss:a_evo}. This is because for $\theta_{\rm sd, i}
        \gtrsim \pi/2$, trajectories are better described as librating about
        CS3. Nevertheless, we can still approximate $A_{\rm i}$ to good
        accuracy as the solid angle enclosed by libration about CS3 with polar
        angle $2\pi + \theta_3 - \theta_2 - \theta_{\rm sd, i}$. This recovers
        the correct $\theta_{\rm f}$-$\p{\theta_{\rm sd, i}}$ mapping for the
        III $\to$ III regime, as can be seen visibly in
        Fig.~\ref{fig:3_ensemble_20_35}.
\end{enumerate}
In summary, starting from an initial condition with phase space area $A_{\rm i}$
at $\eta = \eta_{\rm i} \gg 1$, the five evolutionary tracks are:
\begin{enumerate}
    \item $A_{\rm i} \in \s{0, A_{\rm II}\p{\eta_{\rm \min, III}}}$. Both the II
        $\to$ III and the II $\to$ I tracks are possible.

@@

    \item $A_{\rm i} > \max \p{A_{\rm I} + A_{\rm II}}$. Only the III $\to$ III
        track.
\end{enumerate}
In all cases, the corresponding ranges for $\theta_{\rm sd,i}$ can be computed via
Eq.~\eqref{eq:ai_qsd}. The boundaries between these ranges are overplotted in
Fig.~\ref{fig:ad_ensemble}, where they can be seen to agree well with the
numerical results.
\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{plots_diskdisp/3testo21.png}
